---
title: "EDS223-HW1"
format: html
editor: visual
---

```{r}
#| warning: false
library(tidyverse)
library(sf)
library(here)
library(dplyr)
library(tmap)

# read in geodatabase of EJScreen data at the Census Block Group level
ejscreen <- sf::st_read(here::here("data", "ejscreen","EJSCREEN_2023_BG_StatePct_with_AS_CNMI_GU_VI.gdb")) 
```

```{r}
#| warning: false
# filter to a state you are interested in
california <- ejscreen %>%
  dplyr::filter(ST_ABBREV == "CA") 

# filter to a county you are interested in
ventura <- ejscreen %>%
  dplyr::filter(CNTY_NAME %in% c("Ventura County"))

# find the average values for all variables within counties
california_counties <- aggregate(california, by = list(california$CNTY_NAME), FUN = mean)
# tried 'drop_na()' but mapping didn't work ' 
```

```{r}
# #filter to a state you are interested in
# oregon <- ejscreen %>%
#   dplyr::filter(st_abbrev == "OR")
# 
# # filter to a county you are interested in
# harney <- ejscreen %>%
#   dplyr::filter(cnty_name %in% c("Harney County"))
# 
# #Check it exist
# #Harney <- subset(ejscreen, state_name == 'Oregon' & cnty_name == 'Harney County')
#  
# # find the average values for all variables within counties
# oregon_counties <- aggregate(oregon, by = list(oregon$cnty_name), FUN = mean)
```

# Map 1: Low-income population
```{r}
ventura_low <- tm_shape(ventura) +
  # Using low income varibale
  tm_polygons("LOWINCOME",
              title = "Percent of Low Income Pop.",
              palette = "YlOrRd",
              style = "pretty") +
  
  tm_layout(title = "Ventura County Low Income Population (Pop.)",
            # background color on map
             bg.color = "lightgrey",
            # size of map (portrait, landscape, etc.)
            asp = 1,
            # Bold letters on legend
            legend.title.fontface = 2,
            #Changing font type
            text.fontfamily = "Times New Roman",
            # text size
            legend.text.size = 1)+
  # adding minimap
  tm_minimap()+
    tm_layout(
    attr.position = c("right", "bottom")) +
  tm_compass()+
  tm_scalebar()

print(ventura_low)
```

# Making Map Interactive
```{r}
tmap_mode("view")
```
# Removing Interactive Map. Turning into plot
```{r}
tmap_mode("plot")
```


# Map 2: PM2.5 levels
```{r}
ventura_pm25 <- # Using particulate matter varibale
  tm_shape(ventura)+
  tm_polygons("PM25",
              title = "Particulate Matter 2.5 Per County",
              palette = "Pastel1",
              style = "pretty") +
  
  tm_layout(title = "Particulate Matter 2.5 in Ventura County",
            # background color on map
             bg.color = "lightgrey",
            # size of map (portrait, landscape, etc.)
            asp = 1,
            # Bold letters on legend
            legend.title.fontface = 2,
            #Changing font type
            text.fontfamily = "Times New Roman",
            # text size
            legend.text.size = 1)+
  #tm_minimap()+
  tm_compass()+
  tm_scalebar()
  

ventura_pm25

```

```{r}
#Both Maps Side-by-side
tmap_arrange(ventura_low, ventura_pm25)

```

## For Fun - The State of California Low-Income Population and Air toxics of Cancer Risk in California
- Interesting how most of the central valley and surrounding areas have a higher percent of air toxic cancer risk than anywhere else in the state of California. I have a hypothesis of why that might be, the central valley is known to have lots of agriculture fields. It is one of the highest producers throughout California, the pesticides that the agriculture needs might have an affect towards these communities.
```{r}
# Printing only the state of California
tm_shape(california_counties) +
  # Showcasing Low - income across California
  tm_polygons("LOWINCOME", # Defines how to plot the object
              title = "Percent of Low - Income - Population",
              palette = "pink") +
  tm_symbols(size = "CANCER",
             shape = 8,
             size.legend = tm_legend(orientation = "portrait",
                                     title = "Air toxics Cancer Risk (per County)",
                                     palette = "plasma"))+# Color not working
  #tm_borders() +
  tm_shape(ventura) +
  tm_lines()+
  tm_scalebar(position = c("left", "bottom"))+
  tm_compass(position = c("right", "top")) +
  tm_minimap(position = c("right","top"))
```
